---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(irr)
```

Load Data

```{r}
df <- read.csv("ExampleData1.csv")
df
```

Data in tidy format - convert to table

```{r}
df2 <- df %>% spread(Rater, Expressivity)
df2
```
All raters rate all subjects - fully crossed.

Continuous data - not categorical - not obviously interval based.

```{r}
# drop subject column
df3 <- df2 %>% select(-Subject)
df3
```
Simple per-rater summary stats.

```{r}
summary(df3)
```

Because there are no missing values and the data is continuous - ICC is preferred although K's alpha can be calculated.

Range of rates values is different (especially rater 4).

for ICC:
- two-way (fully crossed)
- prefer consistency over agreement as the rater ranges are different and we do not have enough information to draw further conclusions
- average
- mixed (assuming raters are fixed - not really enough information to determine this)

```{r}
icc(df3, model = "twoway", type = "consistency", unit = "average")
```

ICC(C,k) = 0.445

```{r}
#kripp.alpha(t(as.matrix(df3)), method = "interval")
```
























