i---
title: "SentimentAnalysis"
author: "GaryMcKeown"
date: "07/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(dplyr)
require(tibble)
require(ggplot2)
library(tidytext)
library(plotly)
```

## Exercise 1:
The Combined dataset contains a lingusitic analysis conducted by the LIWC program. LIWC has six emotion 
dimensions: affect, posemo, negemo, anxiety, anger and sad.
The Combined dataset contains headlines from news outlets between the end of 2014 and start of 2015.
The headlines are contained in the variable title and the news outlets are in the variable newsSource.
Work through the commands below and use the headline data in the Combined dataset to create ggplots of the LIWC decisions on the Sentiment.

Import the datasets using RStudio

```{r}
by_newsSource_pos <- Combined %>% 
  group_by(newsSource) %>%
  summarise(meanPos = mean(posemo))

ggplot(by_newsSource_pos, aes(x=newsSource, y=meanPos))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))

by_newsSource_neg <- Combined %>% 
  group_by(newsSource) %>%
  summarise(meanNeg = mean(negemo))

ggplot(by_newsSource_neg, aes(x=newsSource, y=meanNeg))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))
```

Create the new variable polarity by subtracting negemo from posemo.
```{r}
Combined <- Combined %>%
  mutate(polarity = posemo - negemo)

by_newsSource_pol <- Combined %>% 
  group_by(newsSource) %>%
  summarise(meanPol = mean(polarity))

ggplot(by_newsSource_pol, aes(x=newsSource, y=meanPol))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))
```

## Exercise 2 

Do the same thing for the other emotional variables in Combined: affect, anxiety, anger and sad

### Anxiety
```{r}
by_newsSource_anx <- Combined %>% 
  group_by(newsSource) %>%
  summarise(meanAnx = mean(anx))

ggplot(by_newsSource_anx, aes(x=newsSource, y=meanAnx))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))+
  geom_hline(yintercept = 0.25)
```

### Anger
```{r}
by_newsSource_ang <- Combined %>% 
  group_by(newsSource) %>%
  summarise(meanAng = mean(anger))

ggplot(by_newsSource_ang, aes(x=newsSource, y=meanAng))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource))) +
  geom_hline(yintercept = 0.47)
```

### Sadness
```{r}
by_newsSource_sad <- Combined %>% 
  group_by(newsSource) %>%
  summarise(meanSad = mean(sad))

ggplot(by_newsSource_sad, aes(x=newsSource, y=meanSad))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource))) +
  geom_hline(yintercept = 0.29)
```



## Exercise 3 

Put a straight line for the the LIWC text norms in the plot found in Table 3 of the LIWC2015 Development Manual


## Exercise 4 

Move on to look at an inner_join with the Warriner lexicon. This is a collection of emotionally relevant words that has scores for Valence, Arousal and Dominance.

You will have to load the Combined_Tidy and Ratings_WarrinerVAD csv files

```{r}
Combined_Tidy <- Combined_Tidy %>%
  mutate(title=as.character(title))

Combined_Tidy <- as_tibble(Combined_Tidy)

headline_words <- Combined_Tidy %>% 
  unnest_tokens(word, title)

headline_Warriner <- headline_words %>%
  inner_join(Ratings_WarrinerVAD, by = "word")
```

### Valence
```{r}
by_newsSource_val <- headline_Warriner %>% 
  group_by(newsSource) %>%
  summarise(meanVal = mean(Valence))

ggplot(by_newsSource_val, aes(x=newsSource, y=meanVal))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))
```

### Arousal
```{r}

by_newsSource_aro <- headline_Warriner %>% 
  group_by(newsSource) %>%
  summarise(meanAro = mean(Arousal))

ggplot(by_newsSource_aro, aes(x=newsSource, y=meanAro))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))
```

### Dominance -- Also known as control and potency

```{r}
by_newsSource_dom <- headline_Warriner %>% 
  group_by(newsSource) %>%
  summarise(meanDom = mean(Dominance))

ggplot(by_newsSource_dom, aes(x=newsSource, y=meanDom))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))
```

Can you remove the stop words from the headlines and try this again.


```{r}
headline_Warriner <- headline_Warriner %>%
anti_join(stop_words, by = c("word" = "word"))

#valence
by_newsSource_val <- headline_Warriner %>% 
  group_by(newsSource) %>%
  summarise(meanVal = mean(Valence))

ggplot(by_newsSource_val, aes(x=newsSource, y=meanVal))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))

#arousal
by_newsSource_aro <- headline_Warriner %>% 
  group_by(newsSource) %>%
  summarise(meanAro = mean(Arousal))

ggplot(by_newsSource_aro, aes(x=newsSource, y=meanAro))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))

#dominance
by_newsSource_dom <- headline_Warriner %>% 
  group_by(newsSource) %>%
  summarise(meanDom = mean(Dominance))

ggplot(by_newsSource_dom, aes(x=newsSource, y=meanDom))+
  geom_bar(stat = "identity", aes(fill = factor(newsSource)))
```

#Exercise 5 

Combine the variables to create a place for each of them in the 3D VAD space

```{r}
meanVAD <- by_newsSource_val %>%
  left_join(by_newsSource_aro, by="newsSource") %>%
  left_join(by_newsSource_dom, by="newsSource")

plot_ly(meanVAD, x = ~meanVal, y = ~meanAro, z = ~meanDom, color = ~newsSource) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'Valence'),
                      yaxis = list(title = 'Arousal'),
                      zaxis = list(title = 'Dominance')))

scale(meanVAD[,2:4], center = TRUE, scale = TRUE)
```

Perhaps they are better taken to centering around zero, we can use the scale() function to create centralized z-scores of the data moving them to a central origin.

```{r}
scaledMeanVAD <- meanVAD %>%
  mutate(meanVal = scale(meanVal, center = TRUE, scale = TRUE),
         meanAro = scale(meanAro, center = TRUE, scale = TRUE),
         meanDom = scale(meanDom, center = TRUE, scale = TRUE))

plot_ly(scaledMeanVAD, x = ~meanVal, y = ~meanAro, z = ~meanDom, color = ~newsSource) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'Valence'),
                      yaxis = list(title = 'Arousal'),
                      zaxis = list(title = 'Dominance')))
```


#Exercise 6 

Can you apply the scale() function to the individual dimensions and recreate the ggplots in Exercise 4
